---
title: "Line Capacity"
author: "Paul S"
date: "6/6/24"
params: 
  line: "Upfield"
format: 
  html: 
    theme: zephyr
    fig-width: 8
    fig-height: 8
    code-fold: false
    toc: true
    toc-depth: 4
    toc-location: left
    page-layout: full
---



## `r params$line`


### Boarding and Alighting Information

The section below shows how boarding and alighting are distributed across the stations on the `r params$line` line. It is split into four categories: weekdays and weekends, and boarding and alighting (getting off). Each category contains further information about whether the train is going towards flinders street, or away from flinders street.

The y-variable is a percentage of total boardings for that day of week, boarding or alighting and direction, so, each subplot has values which sum to one.

```{r, results='hide', warning=FALSE,message=FALSE}
#| echo: false 

library(tidyverse)

by_line_data = readRDS('../r_objects/by_line_data.Rdata')

upf = by_line_data %>%
  filter(Line_Name == 'Upfield') %>%
  arrange(Station_Chainage) %>% filter(n > 1000)
```

::: {.panel-tabset}

## Weekdays - Boarding

```{r, results='hide', warning=FALSE,message=FALSE}
#| echo: false 

ggplot(upf %>% filter(!isWeekend), mapping = aes(x = reorder(Station_Name, Station_Chainage, decreasing = F), y = pc_boardings) ) +
  geom_col() + scale_x_discrete(guide = guide_axis(angle = 45)) + facet_grid(rows = vars(Direction))


```

## Weekdays - Alighting

```{r, results='hide', warning=FALSE,message=FALSE}
#| echo: false 

ggplot(upf %>% filter(!isWeekend), mapping = aes(x = reorder(Station_Name, Station_Chainage, decreasing = F), y = pc_alightings) ) +
  geom_col() + scale_x_discrete(guide = guide_axis(angle = 45)) + facet_grid(rows = vars(Direction))

```

## Weekends - Boarding

```{r, results='hide', warning=FALSE,message=FALSE}
#| echo: false 

ggplot(upf %>% filter(isWeekend), mapping = aes(x = reorder(Station_Name, Station_Chainage, decreasing = F), y = pc_boardings) ) +
  geom_col() + scale_x_discrete(guide = guide_axis(angle = 45)) + facet_grid(rows = vars(Direction))


```

## Weekends - Alighting

```{r, results='hide', warning=FALSE,message=FALSE}
#| echo: false 


ggplot(upf %>% filter(isWeekend), mapping = aes(x = reorder(Station_Name, Station_Chainage, decreasing = F), y = pc_alightings) ) +
  geom_col() + scale_x_discrete(guide = guide_axis(angle = 45)) + facet_grid(rows = vars(Direction))

```

:::

### Typical Service Profile

::: {.panel-tabset}

## Weekdays

```{r, results='hide', warning=FALSE,message=FALSE}
#| echo: false 
plot <- ggplot(upf %>% filter(!isWeekend), mapping = aes(x = reorder(Station_Name, Station_Chainage, decreasing = T), y = median_departing_load) ) +
  geom_point() + scale_x_discrete(guide = guide_axis(angle = 45))

plot + facet_grid(rows = vars(Direction))

```

## Weekends

```{r, results='hide', warning=FALSE,message=FALSE}
#| echo: false 

plot <- ggplot(upf %>% filter(isWeekend), mapping = aes(x = reorder(Station_Name, Station_Chainage, decreasing = T), y = median_departing_load) ) +
  geom_point() + scale_x_discrete(guide = guide_axis(angle = 45))

plot + facet_grid(rows = vars(Direction))

```

:::

